The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-12 rational.tex
constructions

@@

\paragraph*{Negation distributes over {{+}}. }
One certainly expects that:
{{-(b1 + b2) <-> (-b1) + (-b2)}}
which can be realized as follows:
\begin{center}
  \includegraphics{diagrams/dist_neg_plus.pdf}
\end{center}

\paragraph*{Lifting an operation of positive types to negated types.}
Any operation on types can be applied to the negative versions of these
types. The idea is simply to reverse the flow of values and use the
adjoint of the operation:
\begin{multicols}{2}
Given {{c : b1 <-> b2}}

\begin{center}
  \includegraphics{diagrams/neg_lift.pdf}
\end{center}

>>>>

\paragraph*{Duality preserves the monoidal tensor. }
As with compact closed categories, the dual on the objects distributes
over the tensor. In terms of {{langRevEE}} we have that {{-(b1+b2)}}
can be mapped to {{(-b1)+(-b2)}} and that {{1/(b1*b2)}} can be mapped
to {{(1/b1)*(1/b2)}}. The isomorphism {{-(b1 + b2) <-> (-b1) + (-b2)}}
can be realized as follows:
\begin{center}
  \includegraphics{diagrams/dist_neg_plus.pdf}
\end{center}

The multiplicative construction is similar. This is a distinguishing
difference from *-autonomous categories that are models for Linear
Logic, where the dual does not preserve the tensor but maps to a dual
tensor \cite{curien2009interactive}.

\paragraph*{Duality is a functor.}
Duality in {{langRevEE}} can map objects to their duals and morphisms
to act on dual objects. In other {{c : b1 <-> b2}} to
{{neg~c:-b1<->-b2}} in the additive monoid and to
{{inv~c:1/b1<->1/b2}} in the multiplicative monoid. 
%% Any operation on types can be applied to the negative versions of these
%% types. 
The idea is simply to reverse the flow of values and use the
adjoint of the operation:
\begin{multicols}{2}
\begin{center}
  \includegraphics{diagrams/neg_lift.pdf}
\end{center}

We are able to do this because in {{langRevEE}} every morphism is an
isomorphism and hence has an inverse morphism, whereas in compact
closed categories with no underlying dagger structure we can only
construct an op-functor.

@@
